Online Multiclass Linear Classification with Bandit Feedback

### 2018-12-01 arxiv-version/main.tex
Small fixes

The key observation is that $\Exp[B] = (K-1) \Exp[C]$ since if $S_t =
\emptyset$, there is $1/K$ probability that the algorithm guesses the correct
label and with probability $(K-1)/K$ algorithm's guess is incorrect.
Putting all the information together, we get that
\begin{align*}
\Exp[M]
& = \Exp[A] + \Exp[B] \\
& = \Exp[A] + (K-1) \Exp[C]  \\
& \le (K-1) \Exp[A + C] \\
& = (K-1) \Exp[U]  \; .
\end{align*}

To finish the proof, we need to upper bound the number of updates, $U$. We claim
that $U \le \lfloor (R/\gamma)^2 \rfloor$. The proof of this upper bound is
similar to the proof of the mistake bound for \textsc{Multiclass Perceptron}
algorithm. Let $w_1^*, w_2^*, \dots, w_K^* \in V$ be vectors that satisfy
\eqref{equation:strong-linear-separability-1},
\eqref{equation:strong-linear-separability-2} and
\eqref{equation:strong-linear-separability-3}.
The $K$-tuple $(w_1^{(t)}, w_2^{(t)}, \dots, w_K^{(t)})$
changes only if there is an update in round $t$.
We investigate how $\sum_{i=1}^K \norm{w_i^{(t)}}^2$ and
$\sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}}$ change. If there is an update in round $t$,
we have
\begin{align*}
\sum_{i=1}^K \norm{w_i^{(t+1)}}^2
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}} \norm{w_i^{(t)}}^2 \right) + \norm{w_{\widehat y_t}^{(t+1)}}^2 \\
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}} \norm{w_i^{(t)}}^2 \right) + \norm{w_{\widehat y_t}^{(t)} + (-1)^{z_t} x_t}^2 \\
& = \left( \sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + \norm{x_t}^2 + \underbrace{(-1)^{z_t} 2 \ip{w_{\widehat y_t}^{(t)}}{x_t}}_{\le 0} \\
& \le \left( \sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + \norm{x_t}^2 \\
& \le \left( \sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + R^2 \; .
\end{align*}
Therefore, after $U$ updates,
$$
\sum_{i=1}^K \norm{w_i^{(T+1)}}^2 \le R^2 U \; .
$$
Similarly, if there is an update in round $t$, we have
\begin{align*}
\sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}}
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}} \ip{w_i^*}{w_i^{(t)}} \right) + \ip{w_{\widehat y_t}^*}{w_{\widehat y_t}^{(t+1)}} \\
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}} \ip{w_i^*}{w_i^{(t)}} \right) + \ip{w_{\widehat y_t}^*}{w_{\widehat y_t}^{(t)} + (-1)^{z_t} x_t} \\
& = \left( \sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}} \right) + (-1)^{z_t} \ip{w_{\widehat y_t}^*}{x_t} \\
& \ge \left( \sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}} \right) + \gamma
\end{align*}
where the last inequality follows from \eqref{equation:strong-linear-separability-2}
and \eqref{equation:strong-linear-separability-3} and case analysis whether $z_t = 0$ or $z_t = 1$.
Thus, after $U$ updates,
$$
\sum_{i=1}^K \ip{w_i^*}{w_i^{(T+1)}} \ge \gamma U \; .
$$
Applying Cauchy-Schwartz's inequality twice and assumption
\eqref{equation:strong-linear-separability-1}, we get that
$$
\sum_{i=1}^K \ip{w_i^*}{w_i^{(T+1)}}
\le \sum_{i=1}^K \norm{w_i^*} \cdot \norm{w_i^{(T+1)}}
\le \sqrt{\sum_{i=1}^K \norm{w_i^*}^2} \sqrt{\sum_{i=1}^K \norm{w_i^{(T+1)}}^2}
\le \sqrt{\sum_{i=1}^K \norm{w_i^{(T+1)}}^2} \; .
$$
Combining the upper and lower bounds on $\sum_{i=1}^K \norm{w_i^{(T+1)}}^2$ inequalities, we get
$$
(\gamma U)^2 \le \sum_{i=1}^K \norm{w_i^{(T+1)}}^2 \le R^2 U \; .
$$
We conclude that $U \le (R/\gamma)^2$. Since $U$ is an integer, $U \le \lfloor (R/\gamma)^2 \rfloor$.
\end{proof}

\begin{theorem}[Mistake lower bound]
\label{theorem:strongly-separable-example-mistake-lower-bound}
Let $\gamma$ be a positive real number and let $R$ be a non-negative real
number. For any (possibly randomized) algorithm for the \textsc{Online
Multiclass Classification with Bandit Feedback} with $K \le
\frac{1}{4}(R/\gamma)^2$ classes there exists an inner product space $(V,
\ip{\cdot}{\cdot})$, a non-negative integer $T$ and a sequence of labeled
examples $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)$ in $V \times
\{1,2,\dots,K\}$ that are strongly linearly separable with margin $\gamma$, the
norms satisfy $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$ and the expected
number of mistakes the algorithm makes is at least $\frac{K-1}{2} \lfloor
\frac{1}{16} (R/\gamma)^2 \rfloor$.
\end{theorem}

\begin{proof}
We use probabilistic method. Let $M = \lfloor \frac{1}{16} (R/\gamma)^2 \rfloor$.
Let $V = \R^{M+1}$ equipped with the standard inner product.  Let $e_1, e_2,
\dots, e_{M+1}$ be the standard orthonormal basis of $V$. We define
vectors $v_1, v_2, \dots, v_M \in V$ where $v_i = \frac{R}{\sqrt{2}}(e_i +
e_{M+1})$ for $i=1,2,\dots,M$. Let $\ell_1, \ell_2, \dots, \ell_M$ be chosen
i.i.d. uniformly at random from $\{1,2,\dots,K\}$ and independently of any
randomness used the by the algorithm. Let $T = M K$. We define examples $(x_1,
y_1), (x_2, y_2), \dots, (x_T, y_T)$ as follows:
$$
\begin{array}{lclclclcl}
(x_1, y_1) & = & (x_2, y_2) & = & \cdots & = & (x_K, y_K) & = & (v_1, \ell_1) \; , \\
(x_{K+1},y_{K+1}) & = & (x_{K+2}, y_{K+2}) & = & \cdots & = & (x_{2K}, y_{2K}) & = & (v_2, \ell_2) \; , \\
& \vdots & \\
(x_{(M-1)K+1}, y_{(M-1)K+1})  & = & (x_{(M-1)K+2}, y_{(M-1)K+2}) & = & \cdots & = & (x_{MK}, y_{MK}) & = & (v_M, \ell_M) \; .
\end{array}
$$

@@

The idea behind all the mappings is polynomial approximation. According to the
well known Stone-Weierstrass theorem (see
e.g.~\citep[Section~10.10]{Davidson-Donsig-2010}), on a compact set,
multivariate polynomials uniformly approximate any continuous function. In our
case, we use multivariate polynomial to approximate the indicator function
corresponding to the intersection of $K-1$ halfspaces. This polynomial separates
examples in one class from examples in the other classes. To be able to quantify
$R'$, $\gamma'$ and time complexities of evaluating $\phi(x)$ and $k(x,x')$, we
need to quantify certain parameters of the approximating polynomial (degree,
norm). We construct two different polynomials with different paremeters. The
parameters are quantified in
Theorems~\ref{theorem:polynomial-approximation-1}~and~\ref{theorem:polynomial-approximation-2}
stated below.

@@

\section{Proof of Theorems~\ref{theorem:polynomial-approximation-1}~and~\ref{theorem:polynomial-approximation-2}}
\label{section:proof-of-polynomial-approximation}.

@@

\begin{proof}
To prove the first part, note that $P_n(z)$ and $P_n(-z)$ are non-positive for
$z \in [0,1]$. If $z=1$ then $P_n(z)=P_n(-z) = 0$. For $z \in [0,1)$, we have
$$
\frac{P_n(-z)}{P_n(z)} = \frac{1+x}{1-x} \prod_{i=1}^n \frac{(x+2^i)^2}{(x-2^i)^2} \ge 1 \; .
$$
TODO
\end{proof}

@@

\begin{proof}
Let $M = \sum_{t=1}^T \indicator{\widehat y_t \neq y_t}$ be the number of
mistakes the algorithm makes. Since the $K$-tuple $(w_1^{(t)}, w_2^{(t)}, \dots,
w_K^{(t)})$ changes only if a mistake is made, we can upper bound $\sum_{i=1}^K
\norm{w_i^{(t)}}^2$ in terms of number of mistakes.
If a mistake happens in round $t$,
\begin{align*}
\sum_{i=1}^K \norm{w_i^{(t+1)}}^2
& = \left(\sum_{i \in \{1,2,\dots,K\} \setminus \{y_t, \widehat y_t\} } \norm{w_i^{(t)}}^2 \right) + \norm{w_{y_t}^{(t)} + x_t}^2 + \norm{w_{\widehat y_t}^{(t)} - x_t}^2 \\
& = \left(\sum_{i \in \{1,2,\dots,K\} \setminus \{y_t, \widehat y_t\} } \norm{w_i^{(t)}}^2 \right) + \norm{w_{y_t}^{(t)}}^2 + \norm{w_{\widehat y_t}^{(t)}}^2 + 2 \norm{x_t}^2 + 2 \ip{w_{y_t}^{(t)} - w_{\widehat y_t}^{(t)}}{x_t} \\
& = \left(\sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + 2 \norm{x_t}^2 + 2 \ip{w_{y_t}^{(t)} - w_{\widehat y_t}^{(t)}}{x_t} \\
& \le \left(\sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + 2 \norm{x_t}^2 \\
& \le \left(\sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + 2 R^2 \; .
\end{align*}
So each time a mistake happens, $\sum_{i=1}^K \norm{w_i^{(t)}}^2$ increases by at most $2R^2$. Thus,
$$
\sum_{i=1}^K \norm{w_i^{(T+1)}}^2 \le 2R^2 M \; .
$$
Let $w_1^*, w_2^*, \dots, w_K^* \in V$ be vectors satisfying the
\eqref{equation:linear-separability-1} and
\eqref{equation:linear-separability-2}. We lower bound $\sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}}$. This quantity changes
only when a mistakes happens. If mistake happens in round $t$, we have
\begin{align*}
\sum_{i=1}^K \ip{w_i^*}{w_i^{(t+1)}}
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{y_t, \widehat y_t\}} \ip{w_i^*}{w_i^{(t)}} \right) + \ip{w_{y_t}^*}{w_{y_t}^{(t)} + x_t} + \ip{w_{\widehat y_t}^*}{w_{\widehat y_t}^{(t)} - x_t} \\
& = \left( \sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}} \right) + \ip{w_{y_t}^* - w_{\widehat y_t}^*}{x_t} \\
& \ge  \left( \sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}} \right) + \gamma \; .
\end{align*}
Thus, after $M$ mistakes,
$$
\sum_{i=1}^K \ip{w_i^*}{w_i^{(T+1)}} \ge \gamma M \; .
$$
We upper bound the left hand side by using Cauchy-Schwartz inequality twice and
the condition \eqref{equation:linear-separability-1} on $w_1^*, w_2^*, \dots,
w_K^*$. We have
$$
\sum_{i=1}^K \ip{w_i^*}{w_i^{(T+1)}}
\le \sum_{i=1}^K \norm{w_i^*} \cdot \norm{w_i^{(T+1)}}
\le \sqrt{\sum_{i=1}^K \norm{w_i^*}^2} \sqrt{\sum_{i=1}^K \norm{w_i^{(T+1)}}^2}
\le \sqrt{\sum_{i=1}^K \norm{w_i^{(T+1)}}^2} \; .
$$
Combining all inequalities, we get
$$
(\gamma M)^2 \le \sum_{i=1}^K \norm{w_i^{(T+1)}}^2 \le 2R^2 M \; .
$$
We conclude that $M \le 2(R/\gamma)^2$. Since $M$ is an integer, $M \le \lfloor 2(R/\gamma)^2 \rfloor$.
\end{proof}

>>>>

The key observation is that $\Exp[B] = (K-1) \Exp[C]$. The equality holds
because if $S_t = \emptyset$, there is $1/K$ probability that the algorithm
guesses the correct label and with probability $(K-1)/K$ algorithm's guess is
incorrect. Putting all the information together, we get that
\begin{align*}
\Exp[M]
& = \Exp[A] + \Exp[B] \\
& = \Exp[A] + (K-1) \Exp[C]  \\
& \le (K-1) \Exp[A + C] \\
& = (K-1) \Exp[U]  \; .
\end{align*}

To finish the proof, we need to upper bound the number of updates, $U$. We claim
that $U \le \lfloor (R/\gamma)^2 \rfloor$. The proof of this upper bound is
similar to the proof of the mistake bound for \textsc{Multiclass Perceptron}
algorithm. Let $w_1^*, w_2^*, \dots, w_K^* \in V$ be vectors that satisfy
\eqref{equation:strong-linear-separability-1},
\eqref{equation:strong-linear-separability-2} and
\eqref{equation:strong-linear-separability-3}.
The $K$-tuple $(w_1^{(t)}, w_2^{(t)}, \dots, w_K^{(t)})$
changes only if there is an update in round $t$.
We investigate how $\sum_{i=1}^K \norm{w_i^{(t)}}^2$ and
$\sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}}$ change. If there is an update in round $t$,
we have
\begin{align*}
\sum_{i=1}^K \norm{w_i^{(t+1)}}^2
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}} \norm{w_i^{(t)}}^2 \right) + \norm{w_{\widehat y_t}^{(t+1)}}^2 \\
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}} \norm{w_i^{(t)}}^2 \right) + \norm{w_{\widehat y_t}^{(t)} + (-1)^{z_t} x_t}^2 \\
& = \left( \sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + \norm{x_t}^2 + \underbrace{(-1)^{z_t} 2 \ip{w_{\widehat y_t}^{(t)}}{x_t}}_{\le 0} \\
& \le \left( \sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + \norm{x_t}^2 \\
& \le \left( \sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + R^2 \; .
\end{align*}
Therefore, after $U$ updates,
$$
\sum_{i=1}^K \norm{w_i^{(T+1)}}^2 \le R^2 U \; .
$$
Similarly, if there is an update in round $t$, we have
\begin{align*}
\sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}}
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}} \ip{w_i^*}{w_i^{(t)}} \right) + \ip{w_{\widehat y_t}^*}{w_{\widehat y_t}^{(t+1)}} \\
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{\widehat y_t\}} \ip{w_i^*}{w_i^{(t)}} \right) + \ip{w_{\widehat y_t}^*}{w_{\widehat y_t}^{(t)} + (-1)^{z_t} x_t} \\
& = \left( \sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}} \right) + (-1)^{z_t} \ip{w_{\widehat y_t}^*}{x_t} \\
& \ge \left( \sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}} \right) + \gamma
\end{align*}
where the last inequality follows from \eqref{equation:strong-linear-separability-2}
and \eqref{equation:strong-linear-separability-3} and analysis of the cases $z_t = 0$ and $z_t = 1$.
Thus, after $U$ updates,
$$
\sum_{i=1}^K \ip{w_i^*}{w_i^{(T+1)}} \ge \gamma U \; .
$$
Applying Cauchy-Schwartz's inequality twice and assumption
\eqref{equation:strong-linear-separability-1}, we get that
$$
\sum_{i=1}^K \ip{w_i^*}{w_i^{(T+1)}}
\le \sum_{i=1}^K \norm{w_i^*} \cdot \norm{w_i^{(T+1)}}
\le \sqrt{\sum_{i=1}^K \norm{w_i^*}^2} \sqrt{\sum_{i=1}^K \norm{w_i^{(T+1)}}^2}
\le \sqrt{\sum_{i=1}^K \norm{w_i^{(T+1)}}^2} \; .
$$
Combining the upper and lower bounds on $\sum_{i=1}^K \norm{w_i^{(T+1)}}^2$ inequalities, we get
$$
(\gamma U)^2 \le \sum_{i=1}^K \norm{w_i^{(T+1)}}^2 \le R^2 U \; .
$$
We conclude that $U \le (R/\gamma)^2$. Since $U$ is an integer, $U \le \lfloor (R/\gamma)^2 \rfloor$.
\end{proof}

\begin{theorem}[Mistake lower bound]
\label{theorem:strongly-separable-example-mistake-lower-bound}
Let $\gamma$ be a positive real number and let $R$ be a non-negative real
number. For any (possibly randomized) algorithm for the \textsc{Online
Multiclass Classification with Bandit Feedback} with $K \le
\frac{1}{4}(R/\gamma)^2$ classes there exists an inner product space $(V,
\ip{\cdot}{\cdot})$, a non-negative integer $T$ and a sequence of labeled
examples $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)$ in $V \times
\{1,2,\dots,K\}$ that are strongly linearly separable with margin $\gamma$, the
norms satisfy $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$ and the expected
number of mistakes the algorithm makes is at least $\frac{K-1}{2} \left\lfloor
\frac{1}{16} (R/\gamma)^2 \right\rfloor$.
\end{theorem}

\begin{proof}
We use probabilistic method. Let $M = \left\lfloor \frac{1}{16} (R/\gamma)^2 \right\rfloor$.
Let $V = \R^{M+1}$ equipped with the standard inner product.  Let $e_1, e_2,
\dots, e_{M+1}$ be the standard orthonormal basis of $V$. We define
vectors $v_1, v_2, \dots, v_M \in V$ where $v_i = \frac{R}{\sqrt{2}}(e_i +
e_{M+1})$ for $i=1,2,\dots,M$. Let $\ell_1, \ell_2, \dots, \ell_M$ be chosen
i.i.d. uniformly at random from $\{1,2,\dots,K\}$ and independently of any
randomness used the by the algorithm. Let $T = M K$. We define examples $(x_1,
y_1), (x_2, y_2), \dots, (x_T, y_T)$ as follows:
$$
\begin{array}{lclclclcl}
(x_1, y_1) & = & (x_2, y_2) & = & \cdots & = & (x_K, y_K) & = & (v_1, \ell_1) \; , \\
(x_{K+1},y_{K+1}) & = & (x_{K+2}, y_{K+2}) & = & \cdots & = & (x_{2K}, y_{2K}) & = & (v_2, \ell_2) \; , \\
& \vdots & \\
(x_{(M-1)K+1}, y_{(M-1)K+1})  & = & (x_{(M-1)K+2}, y_{(M-1)K+2}) & = & \cdots & = & (x_{MK}, y_{MK}) & = & (v_M, \ell_M) \; .
\end{array}
$$

@@

The idea behind all the mappings is polynomial approximation. According to the
well known Stone-Weierstrass theorem (see
e.g.~\citep[Section~10.10]{Davidson-Donsig-2010}), on a compact set,
multivariate polynomials uniformly approximate any continuous function.
Intuitively speaking, we use a multivariate polynomial to approximate, on the
unit ball of $\R^d$, the indicator function corresponding to the intersection of
$m=K-1$ halfspaces. Within margin $\gamma$ along the decision boundary, we allow
the polynomial to attain arbitrary value. The polynomial separates examples in
one class from examples in the other classes. To be able to quantify $R'$,
$\gamma'$ and time complexities of evaluating $\phi(x)$ and $k(x,x')$, we need
to quantify certain parameters of the approximating polynomial. We construct two
different polynomials with different paremeters. The parameters are quantified
in
Theorems~\ref{theorem:polynomial-approximation-1}~and~\ref{theorem:polynomial-approximation-2}
stated below.

@@

\section{Proof of Theorems~\ref{theorem:polynomial-approximation-1}~and~\ref{theorem:polynomial-approximation-2}}
\label{section:proof-of-polynomial-approximation}

@@

\begin{proof}
To prove the first part, note that $P_n(z)$ and $P_n(-z)$ are non-positive for
$z \in [0,1]$. If $z=1$ then $P_n(z)=P_n(-z) = 0$. For $z \in [0,1)$, we have
$$
\frac{P_n(-z)}{P_n(z)} = \frac{1+z}{1-z} \prod_{i=1}^n \frac{(z+2^i)^2}{(z-2^i)^2} \ge 1 \; .
$$
TODO
\end{proof}

@@

\begin{proof}
Let $M = \sum_{t=1}^T \indicator{\widehat y_t \neq y_t}$ be the number of
mistakes the algorithm makes. Since the $K$-tuple $(w_1^{(t)}, w_2^{(t)}, \dots,
w_K^{(t)})$ changes only if a mistake is made, we can upper bound $\sum_{i=1}^K
\norm{w_i^{(t)}}^2$ in terms of number of mistakes.
If a mistake happens in round $t$ then
\begin{align*}
\sum_{i=1}^K \norm{w_i^{(t+1)}}^2
& = \left(\sum_{i \in \{1,2,\dots,K\} \setminus \{y_t, \widehat y_t\} } \norm{w_i^{(t)}}^2 \right) + \norm{w_{y_t}^{(t)} + x_t}^2 + \norm{w_{\widehat y_t}^{(t)} - x_t}^2 \\
& = \left(\sum_{i \in \{1,2,\dots,K\} \setminus \{y_t, \widehat y_t\} } \norm{w_i^{(t)}}^2 \right) + \norm{w_{y_t}^{(t)}}^2 + \norm{w_{\widehat y_t}^{(t)}}^2 + 2 \norm{x_t}^2 + 2 \ip{w_{y_t}^{(t)} - w_{\widehat y_t}^{(t)}}{x_t} \\
& = \left(\sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + 2 \norm{x_t}^2 + 2 \ip{w_{y_t}^{(t)} - w_{\widehat y_t}^{(t)}}{x_t} \\
& \le \left(\sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + 2 \norm{x_t}^2 \\
& \le \left(\sum_{i=1}^K \norm{w_i^{(t)}}^2 \right) + 2 R^2 \; .
\end{align*}
So each time a mistake happens, $\sum_{i=1}^K \norm{w_i^{(t)}}^2$ increases by at most $2R^2$. Thus,
$$
\sum_{i=1}^K \norm{w_i^{(T+1)}}^2 \le 2R^2 M \; .
$$
Let $w_1^*, w_2^*, \dots, w_K^* \in V$ be vectors satisfying the
\eqref{equation:linear-separability-1} and
\eqref{equation:linear-separability-2}. We lower bound $\sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}}$. This quantity changes
only when a mistakes happens. If mistake happens in round $t$, we have
\begin{align*}
\sum_{i=1}^K \ip{w_i^*}{w_i^{(t+1)}}
& = \left( \sum_{i \in \{1,2,\dots,K\} \setminus \{y_t, \widehat y_t\}} \ip{w_i^*}{w_i^{(t)}} \right) + \ip{w_{y_t}^*}{w_{y_t}^{(t)} + x_t} + \ip{w_{\widehat y_t}^*}{w_{\widehat y_t}^{(t)} - x_t} \\
& = \left( \sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}} \right) + \ip{w_{y_t}^* - w_{\widehat y_t}^*}{x_t} \\
& \ge  \left( \sum_{i=1}^K \ip{w_i^*}{w_i^{(t)}} \right) + \gamma \; .
\end{align*}
Thus, after $M$ mistakes,
$$
\sum_{i=1}^K \ip{w_i^*}{w_i^{(T+1)}} \ge \gamma M \; .
$$
We upper bound the left hand side by using Cauchy-Schwartz inequality twice and
the condition \eqref{equation:linear-separability-1} on $w_1^*, w_2^*, \dots,
w_K^*$. We have
$$
\sum_{i=1}^K \ip{w_i^*}{w_i^{(T+1)}}
\le \sum_{i=1}^K \norm{w_i^*} \cdot \norm{w_i^{(T+1)}}
\le \sqrt{\sum_{i=1}^K \norm{w_i^*}^2} \sqrt{\sum_{i=1}^K \norm{w_i^{(T+1)}}^2}
\le \sqrt{\sum_{i=1}^K \norm{w_i^{(T+1)}}^2} \; .
$$
Combining all inequalities, we get
$$
(\gamma M)^2 \le \sum_{i=1}^K \norm{w_i^{(T+1)}}^2 \le 2R^2 M \; .
$$
We conclude that $M \le 2(R/\gamma)^2$. Since $M$ is an integer, $M \le \lfloor 2(R/\gamma)^2 \rfloor$.
\end{proof}
